---
title: "QA"
# subtitle: "Data import"
authors: "Michabo Health Science Ltd"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: false
link-citations: true
bibliography: references.bib

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# A note on Quality assessments

> **We will skip in our training today!** Assessing the technical quality of the metabolomics assay is **not** the focus in this tutorial we will not delve any further into quality assessments. However, the trainees should just be aware that the metabolomics data we are using has been assessed at a technical level using quality control samples and passes accetable criteria. 


Whenever using omics datasets, sufficient quality assessments need to be performed to assess the quality of the data from a tehcnical perspective. 

These are specific based on the Omics and technology used. Details for quality assessments in Metabolomics can be found xxx. Details for quality assessments in Transcriptomics can be found xxx.

Quality assessments was performed on all 6 of the participating ring trial labs. All but one passed the quality assessments.

One of the key metrics involved in this assessment is to evaluate intrastudy quality control-samples which are a pooled sample derived from all other samples. The intrastudy QC samples were assessed using Principal Component Analysis (PCA) and assessing the medium relative standard deviation across all features. Following this evaulation the data from RP8 was deemed not be sufficient high quality and so was removed from further analysis.


[@viant2024demonstrating]

```{r table, tidy=FALSE, echo=FALSE}
library(formattable)
library(tidyr)
library(knitr)

data <- read.csv('cefic_matching_quality_assessments.csv', stringsAsFactors = FALSE, header = TRUE, check.names=FALSE)

highlight <- formattable::color_tile("red","red")

formattable(data,caption = '<b>Table 2.1</b>: Quality assessments of intrastudy QCs and biological samples', list(
  area(col = 7, row = 1 ) ~ highlight
  )) 


```


![**Fig 1.1. Cefic MATCHING quality assessments.** Composite of exemplar PCA scores plots (with Hotelling’s T2 ellipse at 95% confidence interval) from each blinded partner for assessment of LC–MS metabolomics data quality after all processing steps, prior to statistical analysis. Plots demonstrate the low technical variability of the intrastudy QC samples (red circles) relative to the larger biological variability of the study samples (treated and control animals, blue crosses) across all partners. The much higher technical variability relative to biological variability observed by partner RP8 (panel f), compared to other partners, is clearly evident. Figure from cefic MATCHING publication"](cefic_matching_quality_assessments.png)



&nbsp;











## References





