---
title: "Differentially abundant molecules - statistical analysis"
#subtitle: "Introductory practical experience in conducting omics-based grouping"
authors: "Michabo Health Science Ltd"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: false
link-citations: true
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Statistical analysis in metabolomics

For the majority of metabolomics studies will use one or more of the following approaches

* Univariate statistics - where each feature is assessed separately (e.g. t-test)
* Multivariate (unsupervised) - where features are assessed together and without the use of training data (e.g. PCA)
* Multivariate (supervised) - where features are assessed together and with the use of labelled training data (e.g. PLS-DA)

In this training we will first use unsupverised multivariate analysis (principal component analysis - "PCA") to assess intensity data and then using univariate analysis (t-tests) to assess the differential abundant molecules compared to the control for each condition.


# PCA of intensity

PCA is unsupervised multivariate statistical approach that is used throughout the analysis of the metabolomics, it allows the high dimensionality of data where we have 1000s of features to be viewed by a smaller number of 'principal components'. These principal components are new variables constructed from linear combinations of the initial features where the first few principal components have the maximum amount of explained variation.

PCA has multiple applications (e.g. to check the quality of the data (i.e. previous section on QCs), to check for outliers, check each step in the data processing (all of which was done with this dataset by Phenome Centre Birmingham and used directly in the interpretation of the biological activity).

Here we will be looking at data that already been checked for data quality (including QC samples, outlier samples) and we are now assessing what we believe is valid bioactivity.



```{r hilic_pos_glog}
library(openxlsx)
# Read in the sample metadata and processed data matrix
sample_metadata <- read.xlsx('female/pmp_HILIC_POS.xlsx', sheet='sample_meta')
glog <- read.xlsx('female/pmp_HILIC_POS.xlsx', sheet='glog')

```



First we need to remove the first column of the matrix which was used to record the feature names
and is not required for this tutorial
```{r hilic_pos_glog_rm}
# Remove the feature names (not required for this analysis)
glog <- glog[,-1]

```

The PCA function within R `prcomp` needs to have the samples as rows and the features
as columns. So next we need transpose the data. 


```{r hilic_pos_glog_transpose}
# transpose 
glog_t <- t(glog)

```

Now let's remove the QC samples as they have already been used for the quality assessments and are no longer required.

```{r hilic_pos_glog_qc_remove}
qc_names <- sample_metadata[,1][sample_metadata$Class=='QC']
sample_metadata_no_qcs <- sample_metadata[sample_metadata$Class!='QC',]

glog_t_no_qcs <- glog_t[!rownames(glog_t) %in% qc_names,]

```

OK... finally ready to perform the PCA

```{r hilic_pos_glog_pca}
pca_no_qcs <- prcomp(glog_t_no_qcs, center = TRUE, scale. = TRUE)
```



We can plot using the `ggfortify` which automatically knows what to plot for PCA outputs from prcomp

```{r hilic_pos_glog_pca_plot}
library(ggfortify)
# the x and y variable can be adjusted to show compare different principal components

autoplot(pca_no_qcs, x=1, y=2, data=sample_metadata_no_qcs, colour="test_substance",  shape="dose_group", frame=TRUE, frame.colour = 'test_substance')+
  scale_colour_manual(values=c("khaki","black","blue", 'red', "grey",'orange', 'magenta', 'yellow', 'green', 'brown', 'purple'))+
  scale_fill_manual(values=c("khaki","black","blue", 'red', "grey",'orange', 'magenta', 'yellow', 'green', 'brown', 'purple'))+
  scale_shape_manual(values=c(4,8,2))+
  theme_bw()

```


> Whilst the above is useful and we can begin to see groups emerging - we are really interested in how similar each test substance compared to the control. So lets perform comparisons to the control for each condition




# Student t-test

Here we will a very common univariate statistical test to use for assessing samples from a specific condition against a control.

Why t-test?



```{r t-test_data}
pqn <- read.xlsx('female/pmp_HILIC_POS.xlsx', sheet='pqn_normalised')

# Label the first column as feature_name so we can access more easily
colnames(pqn)[1] <- 'feature_name'

```

Lets start with doing a single t-test for one feature and one condition

```{r t-test_data_one_feature}
# Get all the control samples (i.e. all those with dose 0)
control_samples <- sample_metadata[,1][sample_metadata$dose_group==0]

# Get the high dose group (2) for test substance 1
ts1_dose_2_samples <- sample_metadata[,1][sample_metadata$dose_group==2 & sample_metadata$test_substance==1]

# Select the 2000th feature which has the name M182T534
M182T534 <- pqn[2000,]

# Perform a t-test for this feature subset based on the chosen samples
ttest_out <- t.test(M182T534[,control_samples], M182T534[,ts1_dose_2_samples])

ttest_out
```

There are two values we will use from the t-test

* **t-statistic (or t-value)**: 
* **p-value**: 


We can use a few loops in function to calculate this for all the conditions within the assay.
Note there are multiple ways to acheive this, for loops are straightforward approach that do not require many additional packages however other packages (like plyr and tidyr) can acheive the same outcome.

```{r t-test_data_one_assay}
ttest_per_assay <- function(pqn, sample_metadata){
    data_long <- melt(ncounts_data, id.vars=c('sample_name'))
  
  
}
```


# Summary of t-tests
Rather than do this on all the assays this has already been performed in the following file







# References
